### 2022-03-28 docs/holodeck-paper/src/ms.tex
manuscript updates

% =================================================================================================
% Holodeck Paper Manuscript
% -------------------------
%
%
%
% =================================================================================================

% general package imports and setup
\input{packages}
% definitions of custom commands
\input{commands}
% journal abbreviations for bibliography
\input{abbrevs}

% =================================================================================================
% ====    Front Matter
% =================================================================================================

\newtotcounter{citnum} % From the package documentation
\def\oldbibitem{} \let\oldbibitem=\bibitem
\def\bibitem{\stepcounter{citnum}\oldbibitem}

\newcommand{\orcidauthorA}{0000-0002-6625-6450} % For author A

\title[Short Title]{Very long and verbose title}
\author[NANOGrav]{the NANOGrav collaboration}
% \author[L.Z.~Kelley et al.]{
%     Luke Zoltan Kelley$^{1}$\thanks{E-mail:lzkelley@northwestern.edu}\orcidA{} \\
%     $^{1}$ \begin{minipage}[t]{\linewidth} Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA), and\\Department of Physics \& Astronomy, Northwestern University, Evanston, IL 60208
%     \end{minipage}
% }

\begin{document}

\maketitle

\begin{abstract}
    abstract
\end{abstract}

\begin{keywords}
    quasars: supermassive black holes
\end{keywords}

% =================================================================================================
% ====    Section 1 - Introduction
% =================================================================================================

\section{Introduction}
    \label{sec:intro}

% =================================================================================================
% ====    Section 2 - Methods
% =================================================================================================

\section{Methods}
    \label{sec:meth}

    \subsection{Source Distributions and Gravitational Wave Calculations}

        Let $N(z)$ denote the number of binaries out to a redshift $z$, corresponding to a comoving volume $V_c$.
        We can calculate the characteristic strain from an ensemble of GW sources as \citep[][Eqs.~5/8]{Phinney-2001},
        	\begin{align}
        	h_c^2(f) = \int_0^\infty \!\! dz \; \frac{d^2 N}{dz \, d\ln f_r} \; h_s^2\lr{f_r}.
        	\end{align}
        In practice, the total number of sources in the Universe is calculated from the number-density, $n \equiv dN / dV_c$.  These can be related as \citep[][Eq.~6]{Sesana+2008},
        	\begin{align}
        	\frac{d^2 N}{dz \, d\ln f_r} = \frac{d n_c}{dz} \frac{dz}{dt} \frac{dt}{d\ln f_r} \frac{d V_c}{dz}.
        	\end{align}
        Note that the total number of binaries in a given frequency interval depends not only on the number density, but also their rate of frequency-evolution.  The faster binaries evolve through a given frequency interval, the fewer binaries will occupy that interval on average.  This is typically described by the binary `hardening' timescale (with respect to frequency), given by the expression,
        \begin{equation}
            \thardf \equiv \diffp{t}{{\ln\!f_r}} = \frac{f_r}{\partial f_r / \partial t}.
        \end{equation}
        Plugging in the relevant cosmographic relationships (see \secref{sec:app_eqs}) we have,
        	\begin{align}
        	h_c^2(f) & = \int_0^\infty \!\! dz \; \frac{dn_c}{dz} \, h^2\lr{f_r} \, 4\pi c \, d_c^2 \lr{1+z} \, \thardf.
        	\end{align}
        The challenge in modeling astrophysical sources of GWs is in calculating their number density.  Numerous approaches are possible, described in the following sections.

    \subsection{Semi-Analytic Modeling}

        Define the distribution function of sources as $f(M,a,z) = d^2 n(M,a,z) / dM da$.  Here $M$ is the total mass of each systems, $a$ is the binary separation, and $z$ is the redshift.  Note that binary frequency $f_r$ could be used almost interchangeably with $a$, and that both evolve as a function of time or equivalently redshift.  This formalism can easily be extended to include other parameters, such as mass-ratio and eccentricity ($q$ and $e$).  We can write the conservation equation for binaries as of function of redshift as,
        \begin{equation}
            \label{eq:conservation}
            \diffp{f}{z} + \diffp{}{m} \lrs{f \diffp{m}{z}} + \diffp{}{a} \lrs{f \diffp{a}{z}} = S_f(m, a, z).
        \end{equation}
        Here $S_f$ is a source/sink function that can account for the creation or destruction of binaries.  Here we assume that all binary `formation' is encapsulated from binaries moving from one part of parameter space (i.e.~large separations and redshifts) to other parts of parameter space (i.e.~smaller separations and redshifts).  Thus we set $S_f = 0$.

        We consider the standard \citep[e.g.~Sesana style; see][]{Chen+2019} semi-analytic model (SAM) formalism of MBH binary populations.  In this style of calculation, $f$ is determined in a region of parameter space that can be observed/estimated, and this is \textit{evolved} to find the distribution in a different region of parameter space that is of interest.  In practice, the observed parameter space is galaxies and galaxy mergers, and the parameter space of interest is close separated MBH binaries that could be GW detectable.  We will express the distribution function as a product of a mass-function, and a pair fraction:
        \begin{equation}
            \label{eq:dist_func}
            f(M,a,z) = \frac{\Phi(M, z)}{M \ln\!10} \cdot F_a(M, a, z),
        \end{equation}
        where the mass-function is, $\Phi(M, z) \equiv \diffp{n_g}{{\logten{M}}}$.
        In \citet{Chen+2019}, for example, the pair fraction is measured over some range of separations, and the separation-dependence is suppressed, i.e.~$f(M, z) = \Phi(M, z) \cdot F(M, z)$ s.t.~$F = \int_{a_0}^{a_1} F_a \, da$.  From $\eqref{eq:conservation}$, we use the chain rule to mix time and redshift evolution, and assume that the mass-change of binaries is negligible, i.e.~$\diffp{m}{t} = 0$, giving:
        \begin{equation}
            \diffp{f}{z} = - \diffp{t}{z} \diffp{}{a} \lrs{f \diffp{a}{t}}.
        \end{equation}
        Plugging-in \eqref{eq:dist_func}:
        \begin{equation}
            \diffp{f}{z} = - \diffp{t}{z} \Phi(M,z) \diffp{}{a} \lrs{F(M,a,z) \diffp{a}{t}}.
        \end{equation}

        The binary population is assumed to be changing only in separation and redshift, which are related by $\partial a / \partial z = (\partial a / \partial t) (\partial t / \partial z)$.  Because the overall number-density is conserved, we can take a finite step in separation and time/redshift, $a\rightarrow a'$ and $z\rightarrow z'$.  Here the time it takes for a binary to go from $a \rightarrow a'$ is $\tau(M,a,z|a')$ which leads to a redshift at the later time of $z' = z'(t + \tau)$.  The standard expression \citep[e.g.~][~Eq.~5]{Chen+2019}, requires making the approximation that,
        \begin{equation}
            \diffp{}{a} \lrs{F(M,a,z) \diffp{a}{t}} \approx \frac{F}{\Delta t} = \frac{F}{\tau(M,a,z|a')}.
        \end{equation}
        Thus giving,
        \begin{equation}
            \diffp{f(M,a',z')}{{z'}} = \diffp{n}{{\log_{10}\!M}{a'}{z'}} = - \diffp{t}{z} \frac{\Phi(M,z) \, F(M,a,z)}{\tau(M,a,z|a')}.
        \end{equation}

        Utimately we are interested in a total number of binaries $N$ such that $dN \equiv n \, dV_c$ for a comoving volume $V_c(z)$.  To find the number of binaries, we use the chain rule to connect the time-evolution of the universe with the evolution of each binary over frequency,
        \begin{equation}
            \diffp{N}{{M}{q}{z}{\ln\!f_r}} = \diffp{n}{{M}{q}{z}} \diffp{z}{t} \diffp{t}{ {\ln\!f_r} } \diffp{V_c}{z}.
        \end{equation}
        Putting this together, we can finally write,
        \begin{equation}
            \label{eq:sam_final}
            \diffp{N}{{M}{q}{z}{\ln\!f_r}} = - \Phi(M,z) \, F(M,a,z) \frac{\thardf}{\tau(M,a,z|a')} \diffp{V_c}{z}.
        \end{equation}

    % % Fig -
    % \begin{figure}
    %     \centering
    %     \includegraphics[width=1\columnwidth]{{{example-image-c}}}
    %     \caption{\textbf{header.} body.}
    %     \label{fig:fig1}
    % \end{figure}
    %
    %
    % \begin{equation}\begin{split}
    %     \label{eq:eq1}
    %     2+2 & = 4 \\
    %         & = 10 - 6
    % \end{split}\end{equation}
    %
    % % Fig -
    % \begin{figure*}
	% 	\centering
    %     \subfloat[first]{{
    %         \includegraphics[width=1.75\columnwidth]{{{example-image-a}}}
    %     }}
	% 	\\
    %     % \qquad
    %     \subfloat[second]{{
    %     	\includegraphics[width=1.75\columnwidth]{{{example-image-b}}}
    %     }}
    %     \caption{\textbf{Head.}  Body.}
    %     \label{fig:fig2}
    % \end{figure*}

\section{Implementation}
    \label{sec:imp}

    The \holodeck{} framework is written primarily in \python{} using a class-based structure for modularity and extensibility.

    \subsection{Semi-Analytic Models}
        \label{sec:imp_sam}

        The core of the semi-analytic modeling framework in \holodeck{} is implementing \eqref{eq:sam_final}.  There are classes for each of the key components: The galaxy stellar-mass function (GSMF, $\Phi$), the galaxy pair fraction \(F\), the merger timescale ($\tau$), and additional a scaling relation for mapping host galaxies via to MBH properties --- typically via an \mmbulge{} relation.  Each component class is implemented generically via basic API methods (e.g.~a GSMF that returns the number density of galaxies at a given stellar-mass and redshift), and can be subclassed to provide a particular implementation (e.g.~an analytic Schechter-function GSMF).  The number (or number density) of binaries is specified over a grid of parameters, typically in a four dimensional space of total mass ($M$), mass ratio ($q$), redshift ($z$), and binary rest-frame orbital frequency ($f_r$).

        The semi-analytic model formalism defines a continuous distribution of MBH binaries.  For many calculations, the discrete nature of MBHB sources is important, not only for examining individual continuous wave sources but also because of the important effects of discrete sources on the shape of the GWB due to the non-linear scaling of GW strain \citep{Sesana+2008}.  Additional, a component of the uncertainty in expected GW signals is due to `cosmic variance', which can be treated as Poisson noise when creating multiple samples (or `realizations') of our binary populations.  To handle discretization and resampling we use the \texttt{kalepy} package \citep{kalepy2021}, in particular, a method of outlier sampling to capture regions of parameter space where finite number effects are important.  The \holodeck{} semi-analytic models specify the number (or number-density) of binaries over a grid, i.e.~$N_{ijkl}$.  By definition, grid cells which have an expected total number of binaries $\gg 1$ will not produce appreciable finite-number effects, so instead they are given a weighting based on a Poisson distribution centered around their expectation value, i.e.~$\mathcal{P}(N_{ijkl})$.  Grid cells with expected numbers of order unity or less do produce important finite-number effects, and thus are directly resampled

% =================================================================================================
% ====    Section 3 - Results
% =================================================================================================
\section{Results}
    \label{sec:res}

    % Fig -
    % \begin{figure*}
    %     \centering
    %     \includegraphics[width=1.5\columnwidth]{{{figs/det-nums_fedd-1.00_voff+3.00_dvel+2.00_tobs+0.70}}}
    %     \caption{\textbf{Distributions of parameters for all binaries (black) and those with a detectable kinematic signature in the secondary AGN (colors).}  Less than $10^{-7}$ of primaries are observable, which are not shown.  Secondary AGN with observable velocity offsets ($\voff$) are shown in blue assuming a sensitivity of \mbox{$\voffsens = 10^3 \kmps$}.  Secondaries with observable changing velocities ($\dvel$) are shown in orange (solid) for a sensitivity $\dvelsens = 10^2 \kmps$, and observing baseline of \mbox{$\tobs = 5\, \yr$}.  Systems where both offsets and changes are detectable (orange, dotted) are mostly a uniform subset of the $\dvel$ sample, and are almost indistinguishable in the 2D contour plots.  An optical flux cutoff of \mbox{$6\E{-14} \mathrm{erg/s/cm}^2$} (AB Mag $\approx 21$) is also imposed, where all AGN are assumed to be accreting at an Eddington fraction of $0.1$.}
    %     \label{fig:det_bins_num}
    % \end{figure*}

% =================================================================================================
% ====    Section 4 - Conclusions
% =================================================================================================
\section{Discussion \& Conclusions}
    \label{sec:disc}

% =================================================================================================
% ====    End Matter
% =================================================================================================

% ====    Acknowledgments

\section*{Acknowledgments}
	I am very thankful to.

    This research made use of \texttt{astropy}, a community-developed core Python package for Astronomy \citep{astropy2013}, in addition to \texttt{Scipy}~\citep{scipy}, \texttt{ipython}~\citep{ipython}, \texttt{jupyter}~notebook~\citep{jupyter}, \texttt{Numpy}~\citep{numpy2011} \& \texttt{SymPy}~\citep{sympy2017}.  All figures were generated using \texttt{matplotlib}~\citep{matplotlib2007}.  Kernel density estimation was performed using the \texttt{kalepy}{} package (\href{https://github.com/lzkelley/kalepy}{github.com/lzkelley/kalepy}) \citep{kalepy2020}.

    The Illustris data is available online at \href{https://www.illustris-project.org/}{www.illustris-project.org} \citep{Nelson+2015}, and Illustris-TNG data at \href{https://www.tng-project.org/}{www.tng-project.org} \citep{Nelson+2019}.

% ====    Bibliography

\let\oldUrl\url
\renewcommand{\url}[1]{\href{#1}{Link}}

\quad{}
\bibliographystyle{mnras}
\bibliography{refs}

\onecolumn
\clearpage

% ====    Appendices

\appendix

    \section{Additional Equations}
        \label{sec:app_eqs}
        A binary is described by its total mass $M$, mass ratio $q \equiv m_2 / m_1$ (s.t.~$q\leq1$), (proper) semi-major axis $a$, and rest-frame orbital frequency $f_r$.  The characteristic gravitational wave mass is the chirp mass, $\mchirp \equiv M \, q^{3/5} \, \lr[-6/5]{1+q}$.

        The GW strain from a circular binary is \citep[][Eq.~7; sky and polarization averaged]{Sesana+2008},
            \begin{equation}
            \label{eq:source_strain}
            h_s = \frac{8}{10^{1/2}} \frac{\left(G\mchirp\right)^{5/3}}{c^4 \, d_L}
                \left(2 \pi f_r \right)^{2/3},
            \end{equation}
        for a luminosity distance $d_L$.  Recall that the luminosity distance and comoving distance are related as $d_L = d_c \, (1+z)$, that the comoving volume of the Universe is given by \citep{Hogg-1999},
        \begin{equation}
            \frac{d V_c}{dz} = 4\pi \frac{c}{H_0} \frac{d_c^2}{E(z)},
        \end{equation}
        and that redshift can be related to time (or age of the universe) as,
        \begin{align}
            \frac{dz}{dt} = H_0 \lr{1+z} E(z).
        \end{align}
        The evolution of the Hubble parameter over redshift can be calculated from the matter-energy constituents of the universe as,
        \begin{align}
            H(z) & = H_0(z) E(z), \\
            E(z) & \equiv \lrs[1/2]{\Omega_R \lr[4]{1+z} + \Omega_m \lr[3]{1+z} + \Omega_k \lr[2]{1+z} + \Omega_\Lambda}.
        \end{align}
        The $\Omega$ terms are the usual redshift-zero fractions of the critical density for radiation, non-relativistic matter, curvature, and dark energy.  We adopt the results of WMAP9 \citep{Hinshaw+2013} for our fiducial cosmology, assuming zero curvature: $H_0 = 69.33 \textrm{ km/s/Mpc}$, $\Omega_m = 0.2880$, $\Omega_\Lambda = 0.7120$.

    \section{Additional Material}
        \label{sec:app}

        \noindent\begin{minipage}{\linewidth}

            % Fig -
            % \centering
            % \includegraphics[width=0.55\columnwidth]{{{figs/doan+2019_jitter}}}
            % \captionof{figure}{Distribution of velocity-offset variations (`jitters') from \citep{Doan+2019}, for the red and blue components of the observed BLs.  Measurements are shown as ticks at $y=0$, and KDE distributions are shown using a Scott's factor bandwidth, and Gaussian kernel.  The purple curve is the KDE distribution taking both red and blue data as independent.  The KDE $50\%$ (median) and $99\%$ (Gaussian standard deviation $\sigma \approx 2.3$) jitter for each component are shown with dashed vertical lines and corresponding labels.}
            % \label{fig:jitter_doan2019}
			%
            % \setlength{\tabcolsep}{8pt}
            % \begin{tabular}{p{1.5cm} c c c c c c c}
            %      & & & $5\%$ & $25\%$ & $50\%$ & $75\%$ & $95\%$ \\ [0.5ex]
            %     \hline
            %     \multirow{4}{*}{All}  & \multirow{4}{*}{-}
            %         & $M \; [M_\odot]$ & $3.4 \times10^{ 6 }$ & $7.2 \times10^{ 6 }$ & $1.8 \times10^{ 7 }$ & $6.0 \times10^{ 7 }$ & $4.1 \times10^{ 8 }$ \\
            %         & & q & $2.9 \times10^{ -2 }$ & $1.7 \times10^{ -1 }$ & $4.0 \times10^{ -1 }$ & $6.8 \times10^{ -1 }$ & $9.3 \times10^{ -1 }$ \\
            %         & & z & 0.24 & 0.41 & 0.53 & 0.62 & 0.69 \\
            %         & & $p \; [\mathrm{yr}]$ & $1.4 \times10^{ 2 }$ & $1.5 \times10^{ 3 }$ & $9.6 \times10^{ 3 }$ & $4.0 \times10^{ 4 }$ & $2.0 \times10^{ 5 }$ \\
            %         & & $a \; [\mathrm{pc}]$ & $2.1 \times10^{ -2 }$ & $1.3 \times10^{ -1 }$ & $4.9 \times10^{ -1 }$ & $1.3 \times10^{ 0 }$ & $4.1 \times10^{ 0 }$ \\
            %         & & $a \; [r_g]$ & $1.2 \times10^{ 4 }$ & $1.4 \times10^{ 5 }$ & $4.5 \times10^{ 5 }$ & $1.2 \times10^{ 6 }$ & $4.1 \times10^{ 6 }$ \\
            %     \hline
            %     \multirow{4}{1.5cm}{Secondary\newline Offset\newline($\voff$)} & \multirow{4}{*}{$0.49\%$}
            %         & $M \; [M_\odot]$ & $1.4 \times10^{ 8 }$ & $4.0 \times10^{ 8 }$ & $7.6 \times10^{ 8 }$ & $1.4 \times10^{ 9 }$ & $3.7 \times10^{ 9 }$ \\
            %         & & q & $9.6 \times10^{ -4 }$ & $4.2 \times10^{ -3 }$ & $1.2 \times10^{ -2 }$ & $3.6 \times10^{ -2 }$ & $1.5 \times10^{ -1 }$ \\
            %         & & z & 0.25 & 0.42 & 0.54 & 0.63 & 0.69 \\
            %         & & $p \; [\mathrm{yr}]$ & $8.3 \times10^{ 2 }$ & $1.5 \times10^{ 3 }$ & $2.3 \times10^{ 3 }$ & $3.7 \times10^{ 3 }$ & $9.0 \times10^{ 3 }$ \\
            %         & & $a \; [\mathrm{pc}]$ & 0.22 & 0.41 & 0.58 & 0.87 & 1.82 \\
            %         & & $a \; [r_g]$ & $5.0 \times10^{ 3 }$ & $1.1 \times10^{ 4 }$ & $1.8 \times10^{ 4 }$ & $2.9 \times10^{ 4 }$ & $4.9 \times10^{ 4 }$ \\
            % \end{tabular}
            % \captionof{table}{\textbf{Parameters of detectable binary systems with redshift $z < 0.7$.}  The indicated quantiles are given for total mass ($M$), mass ratio ($q$), separation ($a$), and orbital period ($p$).  In only four binaries ($2\E{-8}$ of systems) is the primary detectable, and those have parameters: \mbox{$M\approx 2$--$5\E{9} \, \msol$},
            % \hspace{0.5ex} \mbox{$q \approx 0.7$--$0.9$}, \hspace{0.5ex} \mbox{$z \approx 0.6$--$0.8$}, \hspace{0.5ex} \mbox{$a \approx 2.2$--$2.8 \, \pc$}.}
            % \label{tab:obs_pars}

        \end{minipage}

    \twocolumn

\end{document}