### 2023-05-21 docs/holodeck-paper/src/packages.tex

\usepackage{color}
% \usepackage{booktabs}   % Fancier tables
\usepackage{tabularx}
\usepackage{float}
% \usepackage{totcount}
% \usepackage{cuted}
% adds colors to colorspace (https://en.wikibooks.org/wiki/LaTeX/Colors)
\usepackage[dvipsnames, table]{xcolor}
\usepackage{graphicx}
% Force figures to stay in their sections (tex.stackexchange.com/q/279/22806)
\usepackage[section]{placeins}
% \usepackage{subfig}
\usepackage{multirow}
\usepackage{multicol}
% \usepackage{widetext}
% \usepackage{upgreek}    % allow alternative greek letters like $\uptau$
\usepackage{enumitem}   % customize enumerate symbols
\setlist[enumerate]{leftmargin=*}   % Set left-margin of enumerate lists to match the edge
\setlist[itemize]{leftmargin=*}   % Set left-margin of enumerate lists to match the edge
\usepackage[T1]{fontenc}

\usepackage{amsmath}
\usepackage{amssymb}

% `txfonts` : Changes math-greek fonts (I like 'sigma'), NOTE: this must be after ams imports
\usepackage{txfonts}

% easier/convenient derivatives and partial derivatives, for example diffp for partials
\usepackage[thinc]{esdiff}

\usepackage{xfrac}

\usepackage{microtype}   % fix overfull/underfull hbox issues

% For ORCID iDs
% \usepackage{tikz,xcolor,hyperref}
% \definecolor{lime}{HTML}{A6CE39}
% \DeclareRobustCommand{\orcidicon}{
% 	\begin{tikzpicture}
% 	\draw[lime, fill=lime] (0,0)
% 	circle [radius=0.16]
% 	node[white] {{\fontfamily{qag}\selectfont \tiny ID}};
% 	\draw[white, fill=white] (-0.0625,0.095)
% 	circle [radius=0.007];
% 	\end{tikzpicture}
% 	\hspace{-2mm}
% }

% \foreach \x in {A, ..., Z}{\expandafter\xdef\csname orcid\x\endcsname{\noexpand\href{https://orcid.org/\csname orcidauthor\x\endcsname}
% 			{\noexpand\orcidicon}}
% }

\graphicspath{ {figs/} }

% shift around the text on the page so looks good on Letter paper
% \voffset-.6in
% \hoffset0.2in